Analyzing Patterns of Hate Crimes

### 2024-09-26 paper/paper.qmd
second draft, analysis + final check + sketch + reference needed

---
title: "Analyzing Patterns of Hate Crimes: Insights from Toronto's Hate Crime Dataset"
author: 
  - Sakura Hu
thanks: "Code and data are available at: https://open.toronto.ca/dataset/deaths-of-people-experiencing-homelessness/."
date: "2024-09-23"
date-format: long
abstract: "This paper explores the issue of hate crimes in Toronto and their underlying causes, highlighting the urgent need for societal awareness and intervention. Utilizing a dataset of hate crime incidents, we employed statistical methods to analyze the counts attributed to various bias categories. Our findings reveal that the majority of hate crimes are directed toward Jewish, female, and Black individuals, with a concerning trend of increasing occurrences over the years. These results underscore the importance of understanding hate crimes to inform policies and initiatives aimed at fostering inclusivity and reducing discrimination in our communities."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(ggplot2)
library(dplyr)
```

@@

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-yearly-deaths
#| fig-cap: Yearly Distribution of Deaths of People Experiencing Homelessness
#| echo: false

analyze_data <- read_csv("../data/raw_data/raw_data.csv")

```

```{r}
# Reshape the data to long format and count occurrences for each bias
bias_counts <- cleaned_data %>%
  select(RACE_BIAS, ETHNICITY_BIAS, SEXUAL_ORIENTATION_BIAS, GENDER_BIAS, LANGUAGE_BIAS, RELIGION_BIAS, MENTAL_OR_PHYSICAL_DISABILITY) %>%
  pivot_longer(cols = everything(), names_to = "Bias_Type", values_to = "Bias_Value") %>%
  mutate(Value = ifelse(Bias_Value == "None", 0, 1)) %>%
  group_by(Bias_Type) %>%
  summarise(Count = sum(Value))

# Create a bar chart
ggplot(bias_counts, aes(x = reorder(Bias_Type, Count), y = Count, fill = Bias_Type)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "Bias Type", y = "Count of Occurrences", title = "Count of Occurrences Based on Different Bias Types") +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}

# Filter out "None" from the RACE_BIAS column and count occurrences
race_bias_counts <- cleaned_data %>%
  filter(RACE_BIAS != "None") %>%
  count(RACE_BIAS)

# Create the bar chart
ggplot(race_bias_counts, aes(x = reorder(RACE_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Race Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Race Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
# Count occurrences by OCCURRENCE_YEAR without filtering
year_counts <- cleaned_data %>%
  count(OCCURRENCE_YEAR)

# Create the bar chart
ggplot(year_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Year of Occurrence", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Year") +
  scale_x_discrete(drop = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
# Filter out "None" and create a new column for Race Bias categorization
race_counts <- cleaned_data %>%
  filter(RACE_BIAS != "None") %>%
  mutate(RACE_CATEGORY = ifelse(RACE_BIAS == "Black", "Black", "Other"))

@@

# Create the stacked bar chart
ggplot(year_race_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = n, fill = RACE_CATEGORY)) +
  geom_bar(stat = "identity", position = "stack", alpha = 0.8) +
  scale_fill_manual(values = c("Black" = "steelblue", "Other" = "orange")) +
  theme_minimal() +
  labs(x = "Year of Occurrence", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Year and Race Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
# Filter out "None" from the ETHNICITY_BIAS column and count occurrences
ethnicity_bias_counts <- cleaned_data %>%
  filter(ETHNICITY_BIAS != "None") %>%
  count(ETHNICITY_BIAS)

# Create the bar chart
ggplot(ethnicity_bias_counts, aes(x = reorder(ETHNICITY_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "ethnicity Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by ethnicity Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
# Filter out "None" from the LANGUAGE_BIAS column and count occurrences
language_bias_counts <- cleaned_data %>%
  filter(LANGUAGE_BIAS != "None") %>%
  count(LANGUAGE_BIAS)

# Create the bar chart
ggplot(language_bias_counts, aes(x = reorder(LANGUAGE_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Language Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Language Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
# Filter out "None" from the RELIGION_BIAS column and count occurrences
religion_bias_counts <- cleaned_data %>%
  filter(RELIGION_BIAS != "None") %>%
  count(RELIGION_BIAS)

# Create the bar chart
ggplot(religion_bias_counts, aes(x = reorder(RELIGION_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Religion Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Language Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Filter out "None" from the SEXUAL_ORIENTATION_BIAS column and count occurrences
sexual_ori_bias_counts <- cleaned_data %>%
  filter(SEXUAL_ORIENTATION_BIAS != "None") %>%
  count(SEXUAL_ORIENTATION_BIAS)

# Create the bar chart
ggplot(sexual_ori_bias_counts, aes(x = reorder(SEXUAL_ORIENTATION_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Sexual Orientation Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Sexual Orientation Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
# Filter out "None" from the GENDER_BIAScolumn and count occurrences
gender_bias_counts <- cleaned_data %>%
  filter(GENDER_BIAS != "None") %>%
  count(GENDER_BIAS)

# Create the bar chart
ggplot(gender_bias_counts, aes(x = reorder(GENDER_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Gender Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Gender Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@@

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

@@

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

---
title: "Analyzing Patterns of Hate Crimes"
subtitle: "WHAT DID YOU FIND?"
author: 
  - Sakura Hu
thanks: "https://open.toronto.ca/dataset/hate-crimes-open-data/"
date: "2024-09-23"
date-format: long
abstract: "This paper explores the issue of hate crimes in Toronto and their underlying causes, highlighting the urgent need for societal awareness and intervention. Utilizing a dataset of hate crime incidents, we employed statistical methods to analyze the counts attributed to various bias categories. Our findings reveal that the majority of hate crimes are directed toward Jewish, female, and Black individuals, with a concerning trend of increasing occurrences over the years. These results underscore the importance of understanding hate crimes to inform policies and initiatives aimed at fostering inclusivity and reducing discrimination in our communities."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(ggplot2)
library(dplyr)
library(knitr)
```

@@

```{r}
#| echo: false
#| warning: false
#| message: false

cleaned_data <- read_csv("../data/analysis_data/analysis_data.csv")
# Summarize the counts of occurrences for different bias categories by year
bias_summary_by_year <- cleaned_data %>%
  group_by(OCCURRENCE_YEAR) %>%
  summarize(
    race_bias_count = sum(RACE_BIAS != "None"),
    ethnicity_bias_count = sum(ETHNICITY_BIAS != "None"),
    religion_bias_count = sum(RELIGION_BIAS != "None"),
    gender_bias_count = sum(GENDER_BIAS != "None"),
    sexual_orientation_bias_count = sum(SEXUAL_ORIENTATION_BIAS != "None"),
    disability_bias_count = sum(MENTAL_OR_PHYSICAL_DISABILITY != "None")
  )

# Display the summary
bias_summary_by_year %>% 
  kable(
    col.names = c("Year", "Race", "Ethnicity", "Religion", "Gender", "Sexual_Orientation", "Disability"),
    # digits = 1
  )
```

```{r}
#| label: fig-occurence-by-bias
#| fig-cap: Count of Occurrences Based on Different Bias Types
#| echo: false
# Reshape the data to long format and count occurrences for each bias
bias_counts <- cleaned_data %>%
  select(RACE_BIAS, ETHNICITY_BIAS, SEXUAL_ORIENTATION_BIAS, GENDER_BIAS, LANGUAGE_BIAS, RELIGION_BIAS, MENTAL_OR_PHYSICAL_DISABILITY) %>%
  pivot_longer(cols = everything(), names_to = "Bias_Type", values_to = "Bias_Value") %>%
  mutate(Value = ifelse(Bias_Value == "None", 0, 1)) %>%
  group_by(Bias_Type) %>%
  summarise(Count = sum(Value))

# Create a bar chart
ggplot(bias_counts, aes(x = reorder(Bias_Type, Count), y = Count, fill = Bias_Type)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "Bias Type", y = "Count of Occurrences") +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
#| label: fig-yearly-occurrence
#| fig-cap: Occurrences of Hate Crimes by Year
#| echo: false
# Count occurrences by OCCURRENCE_YEAR without filtering
year_counts <- cleaned_data %>%
  count(OCCURRENCE_YEAR)

# Create the bar chart
ggplot(year_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Year of Occurrence", y = "Number of Occurrences") +
  scale_x_discrete(drop = FALSE) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: fig-religion-bias
#| fig-cap: Occurrences of Hate Crimes by Religion Bias
#| echo: false
# Filter out "None" from the RELIGION_BIAS column and count occurrences
religion_bias_counts <- cleaned_data %>%
  filter(RELIGION_BIAS != "None") %>%
  count(RELIGION_BIAS)

# Create the bar chart
ggplot(religion_bias_counts, aes(x = reorder(RELIGION_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Religion Bias Categories", y = "Number of Occurrences") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: fig-Jewish-bias
#| fig-cap: Occurrences of Hate Crimes by Year and Race Bias
#| echo: false
# Filter out "None" and create a new column for Race Bias categorization
race_counts <- cleaned_data %>%
  filter(RELIGION_BIAS != "None") %>%
  mutate(RELIGION_CATEGORY = ifelse(RELIGION_BIAS == "Jewish", "Jewish", "Other"))

# Count occurrences by OCCURRENCE_YEAR and RACE_CATEGORY
year_religion_counts <- race_counts %>%
  count(OCCURRENCE_YEAR, RELIGION_CATEGORY)

# Calculate percentage for each category within each year
year_religion_counts <- year_religion_counts %>%
  group_by(OCCURRENCE_YEAR) %>%
  mutate(percentage = n / sum(n) * 100)

# Create the stacked bar chart
ggplot(year_religion_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = percentage, fill = RELIGION_CATEGORY)) +
  geom_bar(stat = "identity", position = "stack", alpha = 0.8) +
  scale_fill_manual(values = c("Jewish" = "steelblue", "Other" = "orange")) +
  theme_minimal() +
  labs(x = "Year of Occurrence", y = "Number of Occurrences") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(label = scales::percent(percentage / 100)), 
            position = position_stack(vjust = 0.5), 
            color = "white", size = 4)  # Adjust size and color as needed
```

```{r}
#| echo: false
# Filter out "None" from the RACE_BIAS column and count occurrences
race_bias_counts <- cleaned_data %>%
  filter(RACE_BIAS != "None") %>%
  count(RACE_BIAS)

# Create the bar chart
ggplot(race_bias_counts, aes(x = reorder(RACE_BIAS, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.8) +
  theme_minimal() +
  labs(x = "Race Bias Categories", y = "Number of Occurrences", 
       title = "Occurrences of Hate Crimes by Race Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| echo: false
# Filter out "None" and create a new column for Race Bias categorization
race_counts <- cleaned_data %>%
  filter(RACE_BIAS != "None") %>%
  mutate(RACE_CATEGORY = ifelse(RACE_BIAS == "Black", "Black", "Other"))

@@

# Calculate percentage for each category within each year
year_race_counts <- year_race_counts %>%
  group_by(OCCURRENCE_YEAR) %>%
  mutate(percentage = n / sum(n) * 100)

# Create the stacked bar chart with percentage labels
ggplot(year_race_counts, aes(x = as.factor(OCCURRENCE_YEAR), y = percentage, fill = RACE_CATEGORY)) +
  geom_bar(stat = "identity", position = "stack", alpha = 0.8) +
  scale_fill_manual(values = c("Black" = "steelblue", "Other" = "orange")) +
  theme_minimal() +
  labs(x = "Year of Occurrence", y = "Percentage of Occurrences", 
       title = "Percentage of Hate Crimes by Year and Race Bias") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(aes(label = scales::percent(percentage / 100)), 
            position = position_stack(vjust = 0.5), 
            color = "white", size = 4)  # Adjust size and color as needed
```

@@

@@

@@

@@
